Mental Health of U.S. Citizens During the COVID-19 Pandemic

### 2022-03-21 outputs/paper/paper.Rmd
final changes

Below contains a summary of the variables of interest using vtable[@vtable] and kableExtra[@kableExtra]:

```{r fig0, warning=FALSE, fig.cap="Summary of US GSS Variables", message = FALSE, echo=FALSE, fig.width=5, fig.height=5}
gss2021 <- gss2021 %>% 
  mutate(`Mental Health Score` = hlthmntl) %>%
  mutate(`Physical Health Score` = hlthphys) %>%
  mutate(`Age when first child born` = agekdbrn)
USData <- gss2021 %>% select(`Mental Health Score`, `Physical Health Score`, age, `Age when first child born` )
sumtable(USData,  title = 'US GSS Summary Statistics', out = 'kable')
```

@@

Covid-19 undoubtedly had a significant impact on the 2021 GSS Survey. With the cancellation of in-person interviews and the shift onto web and phone mode, this year’s GSS interviewers were unable to interact with the respondents; skipping questions, and volunteer responses were allowed. This may introduce within the responses without any unbiased medium, such as the interviewer to assess. Furthermore, as mentioned in the #Bias Section, the randomness of the survey has shifted and the response rate drastically decreased due to the impact of the pandemic. Although weighting has been introduced in an attempt to alleviate the impacts, these are still factors to consider.

>>>>

```{r tab1, warning=FALSE, fig.cap="Summary of US GSS Variables", message = FALSE, echo=FALSE, fig.width=5, fig.height=5}
gss2021 <- gss2021 %>% 
  mutate(`Mental Health Score` = hlthmntl) %>%
  mutate(`Physical Health Score` = hlthphys) %>%
  mutate(`Age when first child born` = agekdbrn)
USData <- gss2021 %>% select(`Mental Health Score`, `Physical Health Score`, age, `Age when first child born` )
sumtable(USData,  title = 'US GSS Summary Statistics', out = 'kable')
```

Table \@ref(tab:tab1) contains a summary of the variables of interest using vtable[@vtable] and kableExtra[@kableExtra].

From this table, we observe the average mental health of respondents as well as their physical health as below the average of the possible responses (3 being the statistically average response), with a slightly higher average physical health. Their standard deviation, or measure of variation; is nearly identical. We also observe the summary statistics of the age the respondents were when they had their first child.

@@

Covid-19 undoubtedly had a significant impact on the 2021 GSS Survey. With the cancellation of in-person interviews and the shift onto web and phone mode, this year’s GSS interviewers were unable to interact with the respondents; skipping questions, and volunteer responses were allowed. This may introduce within the responses without any unbiased medium, such as the interviewer to assess. Furthermore, as mentioned in the [Bias][Bias] Section, the randomness of the survey has shifted and the response rate drastically decreased due to the impact of the pandemic. Although weighting has been introduced in an attempt to alleviate the impacts, these are still factors to consider.
